### 2019-08-20 Writeup/Thesis/chapters/lit_review.tex
Changes; - Added some changes from supervisor - Began fixing structure of the document

An E2EE connection uses the key types discussed previously to exchange messages between verified parties securely. However, this hinges on the verification of the initial parties, if one party impersonates another and receives sensitive communication, the secure encryption is completely circumvented. Therefore, the correct identification of parties is crucial to maintaining security. The exploitation of this is known as a Man-in-the-middle (MiTM) attack. This attack is commonly bidirectional, and if executed correctly, there is often no discerning change to the user experience. Figure \ref{fig:mitm} contains a visual representation of this attack.

\begin{center}
    \input{diagrams/MITM/basic}
    \begin{figure}[h]
        \caption{Photo depicting a MITM attack}
        \label{fig:mitm}
    \end{figure}
\end{center}

Due to the assumption that the attacking party (Eve) does not have the same public-private key pair as either party (Alice or Bob), the fingerprint of the public keys are, therefore, different, and, thus can be used for identification. There are, however, issues in how a user should link a key's fingerprint to a real-world entity. This paper will, however, assume that users have access to the respective parties' correct fingerprint, as the protocols used to achieve this are outside the project's scope.

The fingerprint of the key is generated by running the main key components through a secure one-way hash function such as SHA-1. This process produces a digest of a fixed length that can be used to compare keys. Therefore, the comparison of expected and actual fingerprints can be used to detect MiTM attacks.
Historically these have been represented as a hexadecimal string whereon verification fingerprints are compared between two substrates, for example, a monitor screen and a business card.
This process is known as the \textit{``authentication ceremony.''} System designers are, however, moving away from this structure and are now implementing a combined key. This change is due to the assumed increase in usability. Users now compare a single fingerprint, instead of two sets separately. In the case of WhatsApp for example, the connection fingerprint of two parties keys is the first 30 bytes of an SHA-512\footnote{5200 iterations} hash of each parties' identity key, these are then both concatenated together to form a single key\cite{whatsapp2017paper}. This process produces a unique-key communication pair.
\\\\
Due to the manual comparison of the fingerprint, length, and format is a core design consideration. Prior research has shown that the average human can only hold around 7-digits worth of data in their working memory\cite{miller1956magical}. This rules out the possibility of comparing complete digests. For example, SHA-1 is 40 hex digits (160-bit) and, therefore, difficult to effectively compare. Therefore, if human interaction is required; there is a need for schemes that work effectively with consideration to individual limitations.

Research in this area has taken various encoding schemes and compared their fallibility to impersonated attacks. The main elements of comparison have been the ``accuracy of attack detection" and ``time to compare." In this paper, these are considered the metrics of ``security'' and ``usability,'' respectively.

\section{Definition of terms}
Throughout the paper, various terms are used to define elements of the authentication ceremony relevant to fingerprints verification. Explicit descriptions are, therefore, stated in this section. These are used to remove ambiguity and underlying associations readers may have for the words chosen.
\\\\
\textbf{\textit{Encoding schemes}} - are the physical method of encoding used to represent the fingerprint. For example, ``Hexadecimal'' or ``Words'' are a unique way to represent a fingerprint and thus, are encoding schemes.
\\\\
\textbf{\textit{Method of comparison}} - is the way a user assesses the encoding scheme. The most common example and the one used as default is ``Compare-and-Confirm'' (CaC). CaC, as the name suggests, is the \textit{comparison} of two fingerprints on different devices or mediums that are then \textit{confirmed}. The following sections explain alternative methods of comparisons.

@@

\subsubsection*{Encoding scheme performance}
Results 
from the literature consistently show the effectiveness of language-based encodings such as Words or Sentences with accuracies ranging up from 94\% \cite{dechand2016empirical}\cite{tan2017can}\cite{kainda2009usability}. In all cases, these were the best schemes from the sets assessed. The exception to this is the work performed by \textbf{Hsiao \textit{et al.}}\cite{hsiao2009study} in 2009 with Words achieving an abnormal accuracy of 63\%.
\\\\
Aside from textual representations were graphical schemes. Examples of schemes assessed were: Random Art\cite{perrig1999hash}, Flag\cite{ellison2003public}, T-Flag\cite{lin2010spate}, Vash\footnote{https://github.com/thevash/vash}, OpenSSH visual host key and Unicorns\footnote{https://unicornify.pictures/}, among others. These schemes had mixed accuracy with ranges as large as 50\% - 94\% in work by \textbf{Hsiao \textit{et al.}}\cite{hsiao2009study}. The only other paper assessing graphical representations was the work of \textbf{Tan \textit{et al.}}\cite{tan2017can} where they also achieved mixed results with accuracies ranging from 46\% to 90\%.
\\
\begin{table}[h!]
    \makebox[\textwidth][c]{
        \input{tables/fingerprint_times.tex}
    }%
    \caption{Timing results in seconds for the related schemes}
    \label{tab:times}
\end{table}
\\
In terms of usability of graphical schemes, the literature concurred on their high usability. The comparison speed of these schemes were all among the quickest (See Table \ref{tab:times} for an overview of timings). Other work also indirectly agreed with graphical encodings having significantly quicker comparison times compared to non-graphical schemes \cite{dechand2016empirical}\cite{kainda2009usability}.
In terms of research into the performance of graphical schemes, the literature does not contain an extensive review with literature only containing two papers. There is also no overlap in the schemes assessed with each paper reviewing a unique set. This area is, therefore, a promising candidate for further research.

One unique paper in this research area was the work by \textbf{M. Shirvanian \textit{et al.}}\cite{shirvanian2017pitfalls} produced in the context of secure messaging pairing. This paper was unique for several reasons. First was to consideration for ``remote-vs-proximity'' pairing where this is the first consideration of this aspect found in the literature. There is room for further research to compare encoding schemes in the context of ``remote-vs-proximity.'' Another unique aspect was the end-to-end encryption context of the study.

@@

Alongside this, it was shown in work by \textbf{Hsiao\textit{et al.}}\cite{hsiao2009study} that age and gender do not affect the accuracy of the scheme. However, younger participants were considerably faster. Furthermore, findings also showed that language comprehension helped in discerning small differences between schemes encoded in Chinese, Japanese, Korean, or English. Subsequently, knowledge of the language did not assist in differentiating more significant changes in the schemes as these had high accuracy regardless. These were unique considerations. Further work could aim to corroborate these conclusions.

\begin{table}[h!]
    \makebox[\textwidth][c]{
        \input{tables/fingerprint_comparison.tex}
    }%
    \caption{Accuracy of correct comparison for the encoding schemes assessed}
    \label{tab:results}
\end{table}

Tables \ref{tab:times} \& \ref{tab:results} contain the accuracy results of all papers assessed; this is to aid in visual comparison. Each paper used a different metric for measuring accuracy. Therefore, all results have been translated into ``overall accuracy.''

@@

The starkest limitations of the literature are the range of participants and encoding entropy. The worst studies tested only 22-bits of entropy. This lack of entropy makes it challenging to compare results directly. One of the papers this most effects is the early work by \textbf{Hsiao \textit{et al.}}\cite{hsiao2009study}  where their highest entropy is 28-bits. This level of entropy was inadequate even at the time of publication. There is an attempt by the authors' to address this issue in the later stages of the paper where they write: \textit{``[...] increasing entropy is not a solution because it sacrifices usability and accuracy. With more entropy, representations will contain longer sequences of characters or more minute details which will lead to increased time and errors during comparisons.''}. This statement is backed up with no empirical evidence, and the authors fail to consider how the low entropy would affect the overall security of the schemes.

@@

One consideration of note is that made by Dechand \textit{et al.} \cite{dechand2016empirical} where they briefly consider medical conditions such as ADHD and reading or visual disorders and the way they affect the comparison's effectiveness. The results highlight a slight reduction in overall accuracy, however, due to their small sample size, the results cannot be considered conclusive. This aspect is unique to all literature assessed. Further work would be required to produce conclusive results. Therefore, highlighting a gap in the literature.

One glaring issue with the demographical health of \textbf{Ersin Uzun \textit{et al.}}\cite{uzun2007usability} study is the use of two entirely different groups of participants. Not only were their demographics different, but they were from different countries (America and Finland). Different cultures contain inherent biases and assumptions. Changes were made pragmatically regarding the results from the first round of 40 participants. These were then re-assessed, alongside the direct comparison of results. This aspect casts vast doubts on the validity of the results with no consideration made by the authors to control external factors that may affect the performance of the method of comparison.

Hexadecimal and numerical schemes were not included as encoding schemes in work by \textbf{Hsiao \textit{et al.}}\cite{hsiao2009study} (some of the most widespread encoding schemes). The justification was the schemes similarities to Base32  alongside "well-known deficiencies". This point was provided with no further justification or quantified. Moreover, it is also inconsistent with available research, for example, in \textit{``Empirical Study of Textual Key-Fingerprint Representations''}\cite{dechand2016empirical} it was shown numerical representations performed significantly better than that of Base32.

\begin{table}[h!]
    \makebox[\textwidth][c]{
        \input{tables/fingerprint_attacker_stren.tex}
    }%
    \caption{Paper attribute comparison}
    \label{tab:attacker}
\end{table}

@@

\subsubsection*{Topic conclusion}
Overall, this review has identified several key areas suitable for further work.
The first is the performance assessment of graphical encoding schemes. Recreation of pre-existing results from \cite{hsiao2009study}\cite{tan2017can} is required to corroborate current conclusions and validate results.
Another gap in the research is the consideration into the utilisation of encoding schemes in realistic conditions, i.e. ``remove vs proximity." This topic was initially covered by \cite{shirvanian2017pitfalls}, but their scope was limited. Further work, therefore, could increase the scope and touch upon a large number of schemes in these settings. The final aspect for further work is the limited consideration into the feasibility of attacks on encoding schemes. All of the papers assessed simulated attacks and had minimal consideration for the execution of these attacks. Therefore, further work could delve
into the implementation and feasibility of such attacks.

@@

Some of the oldest preliminary work into visual encoding schemes is by \textbf{Adrian Perrig \textit{et al.}}\cite{perrig1999hash} in the creation of their scheme ``Random Art" in 1999. The motivation for creating such a scheme was the perceived flaws in the ways humans verify and compare written information. As mentioned in previous sections, visual encoding schemes have been shown to have mixed success, with low security being one of their most alarming flaws. This research laid the foundation for further work in analysing the security of visual encoding schemes.

Further research into the creation of unique visual hash schemes has been performed by \textbf{C Ellison \textit{et al.}} \cite{ellison2003public} (Flag), \textbf{Yue-Hsun Lin \textit{et al.}}\cite{lin2010spate} (T-Flag) and work by \textbf{M.  Olembo \textit{et al.}}\cite{olembo2013developing}. Each publication has provided a new way  to represent a key fingerprint visually. Alongside the academic literature, there are informally presented methods of visual fingerprints such as Unicorns\footnote{https://unicornify.pictures/} and Robots\footnote{https://github.com/e1ven/Robohash}. This list is by no means exhaustive but is used to depict the amount of research and work invested into graphical hash representations.

@@

Overall the paper provides a foundation for formalising the creation of effective wordlists. A limitation is the lack of empirical data gathered on the performance. However, this was later evaluated in work by Dechand \textit{et al.} \cite{dechand2016empirical} and shown to be an effective encoding scheme.

Other research of note is work by \textbf{M. Goodrich \textit{et al.}}\cite{goodrich2006loud} called \textit{Loud and Clear: Human-Verifiable Authentication Based on Audio}. As the name suggests, the authors were researching ways to improve current methods of secure device pairing. The unique aspect of this work is the use of a Text-to-Speech system reading out syntactically correct English sentences. MadLibs\footnote{https://en.wikipedia.org/wiki/Mad\_Libs} was the base for these sentences. MadLibs is where static placeholders are replaced with a list potential words.\\
This work is an extension to work performed by Juola and Zimmermann\cite{juola1996whole} as they aimed to emulate the techniques used in PGPfone. The paper's findings are limited by the lack of empirical data backing up claims made by the author the performance of the system, and security is only theoretically assessed.

Aside from this research, there have been further informal implementations of fingerprint encodings. The first being by \textbf{Michael Rogers}\footnote{https://github.com/akwizgran/basic-english}. Rogers' implementation is a program designed to map fingerprints to pseudo-random poems. This implementation was again, empirically evaluated by Dechand \textit{et al.}\cite{dechand2016empirical}. Ealier work by \textbf{N. Haller} with the S/KEY\cite{haller1995s} shows the implementation of a system designed to represent a hash as a series of six short words. However, this system is designed for a one-time-password purpose and only provides word mappings for basic human usability of the password and not within a fingerprint verification context. Therefore, the wordlist has not been designed with pronounceability in mind.

@@

\subsubsection*{Topic conclusion}

In conclusion to this topic, the current research has primarily focused on the research and creation of visual representations. Research for textual fingerprints is fragmented and incomplete with work Juola and Zimmermann 
\cite{juola1996whole} and M. Goodrich \textit{et al.}\cite{goodrich2006loud} providing meaningful research to build upon in terms of word and sentence based encodings. The fragmentation of this research leaves room for further work into this topic area. Alongside this, findings from the previous section's research show that human language based encodings provided the best usability and, therefore, should be a target for further research looking to improve upon their security and usability.

@@

The only other relevant research on this topic is the work by \textbf{M Shirvanian \textit{et al.}}\cite{shirvanian2014wiretapping} 
and their paper: \textit{``Wiretapping via Mimicry: Short 
Voice Imitation Man-in-the-Middle Attacks on Crypto 
Phones''}. Further research in the area of ``human voice impersonation'' has received lots of attention \cite{mukhopadhyay2015all}\cite{chen2017you}\cite{wu2015spoofing}. This paper was chosen over other alternatives due to its specific use of encoding schemes in its evaluation.

In this paper, the authors develop a way to 
impersonate users when authenticating 
Short-authentication-Strings (SAS) in the pairing of 
Crypto-phones. To achieve this impersonating they propose 
two methods: ``Short voice re-ordering attack'' where an 
arbitrary SAS string is recreated by re-ordering snipped 
obtained from eavesdropping a previous connection
and ``Short voice morphing attacks'' whereby the use of 
previously eavesdropped audio snippets the attacker can
morph their voice to match that of the victim. With 
these methods, they aimed to attack encodings of Numbers, 
PGP word list (previously discussed work by Juola and 
Zimmermann \cite{juola1996whole}) and MadLib (M. Goodrich 
\textit{et al.}\cite{goodrich2006loud} work also 
previously discussed). The effectiveness of these attacks 
was evaluated with a study involving 30 participants.

@@

\section{Overall Summary}
In summary of the literature gaps areas discovered; encoding scheme performance requires further work in the performance of all graphical schemes to back up the results made by other work. Furthermore, there is a gap with the assessment of encoding schemes in the context of "remote-vs-proximity" first proposed by M. Shirvanian et al. as it would arguably provide a better simulation of real-world scenarios.

In the context of human demographics and the way they alter the performance of the schemes; further research is required into how the fluency of languages affects authentication ceremony. This further work would be a continuation of the initial work by Hsiao et al. and could allow the creation of schemes that are better suited to specific languages.
Alongside this, there is a research gap in the investigation of how mental impairments affect the performance. This aspect is useful due to the number of potential users being affected due to systems not being designed with them in mind. For example, 7\% of the population have been identified as having dyslexic tendencies\cite{peterson2012developmental}. This fact means that with a UK population of around 63,000,000\footnote{From the 2011 Census collated by the Office for National Statistics}  4,410,000 people could benefit from encoding schemes designed to work well with dyslexia. Dyslexia, however, is only one of many impairments that could affect a scheme's performance, thus further contributing to the substantial effect further work could have on this area.

>>>>

An E2EE connection uses the key types discussed previously to exchange messages between verified parties securely. However, this hinges on the verification of the initial parties, if one party impersonates another and receives sensitive communication, the secure encryption is completely circumvented. Therefore, the correct identification of parties is crucial to maintaining security. The exploitation of this is known as a Man-in-the-middle (MiTM) attack. This attack is commonly bidirectional, and if executed correctly, there is often no discerning change to the user experience. Figure \ref{fig:mitm} contains a visual representation of this attack. In the diagram Eve (attacker) is impersonating Alice to Bob and Bob to Alice. Therefore, as both directions of communication are compromised all traffic will route through the attacker Eve.

\begin{center}
    \input{diagrams/mitm.tex}
    \begin{figure}[h]
        \caption{Photo depicting a MITM attack}
        \label{fig:mitm}
    \end{figure}
\end{center}

Due to the assumption that the attacking party (Eve) does not have the same public-private key pair as either party (Alice or Bob), the unique aspects of the keys can be used to identify parties. This is achieved succinctly through the use of a fingerprint. A fingerprint is a string of alphanumerical characters that are unique to each key. Therefore the fingerprint of the two respective public keys are different, and, thus can be used for identification. There are, however, issues in how a user should link a key's fingerprint to an identity. One example way to achieve this is by meeting in-person to exchange fingerprints. This paper will, however, assume that users have been through this process.

The fingerprint of the key is generated by running the main key components through a secure one-way hash function such as SHA-256. This process produces a digest of a fixed length that can be used to compare keys. Therefore, the comparison of expected and actual fingerprints can be used to detect MiTM attacks.

Historically fingerprints have been represented as a hexadecimal string, whereon verification fingerprints are compared between two substrates, for example, a monitor screen and a business card. Figure \ref{fig:businessCard} shows an example business card. A user will then use the fingerprint present on the card to compare to the one available on a device. This will have to be performed by each user. This process is known as the \textit{``authentication ceremony.''}

\begin{figure}[h!]
    \centering
    \fbox{
        \input{diagrams/business_card.tex}
    }
    \caption{Example business card with fingerprint}
    \label{fig:businessCard}
\end{figure}

System designers are, however, moving away from the structure of one fingerprint per identity and are now implementing a combined key. This change is due to the assumed increase in usability. Users now compare a single fingerprint together, instead of two sets separately. In the case of WhatsApp for example, the connection fingerprint of two parties keys is the first 30 bytes of an SHA-512\footnote{5200 iterations} hash of each parties' identity key, these are then both concatenated together to form a single fingerprint\cite{whatsapp2017paper}. This process produces a unique key per communication pair.

Due to the manual comparison of the fingerprint, length, and format is a core design consideration. Prior research has shown that the average human can only hold around 7-digits worth of data in their working memory\cite{miller1956magical}. This shows that comparing complete digests would be cumbersome and error-prone. For example, SHA-1 is 40 hex digits (160-bit) and, therefore, difficult to effectively compare. Therefore, if human interaction is required; there is a need for schemes that work effectively with consideration to human limitations.

\section{Encoding schemes}
\label{sec:encodingSchemes}
Encoding schemes are the physical method of encoding used to represent the fingerprint. The most common way to represent the fingerprint is alphanumerically with either \textbf{Hexadecimal} (a-f/0-9) or \textbf{Base32} (2-7/A-Z). These encoding schemes are the most popular due to their intuitive simplicity and lack of unique hardware requirements such as colour screens.

Fingerprints can also be encoded using natural language. The fingerprint can be chunked and mapped to a set or words. The same principle can be used to fill placeholders in a pre-defined sentence allowing the simple formation of syntactically accurate English sentences. Moreover, other languages can be used such as Chinese, Japanese or Korean to map fingerprint chunks to characters.

\begin{table}[h!]
    \centering
    \input{tables/background/encoding_examples.tex}
    \caption{Examples for text based encodings}
\end{table}

Graphical schemes can also be used to represent a fingerprint.

\textbf{Random Art}. Random Art engine\cite{perrig1999hash} takes any length of input and processes it via SHA-1 to produce a digest that is then converted into an image. Figure \ref{fig:randomArt} contains three examples from the Random Art Gallery\footnote{\url{http://www.random-art.org/}}.

\begin{figure}[h!]
    \centering
    \minipage{0.32\textwidth}
        \includegraphics[width=.9\textwidth]{random_art/random_art1.jpg}
    \endminipage
    \minipage{0.32\textwidth}
        \includegraphics[width=.9\textwidth]{random_art/random_art2.jpg}
    \endminipage
    \minipage{0.32\textwidth}
        \includegraphics[width=.9\textwidth]{random_art/random_art3.jpg}
    \endminipage
    \caption{Random Art examples}
    \label{fig:randomArt}
\end{figure}

\textbf{Flag}. Flag is a visual hash scheme proposed by \textbf{C. Ellison} and \textbf{S. Dohrmann}\cite{ellison2003public}. Flag consists of four coloured strips with $2^n$ number of possible colours in each strip.

\textbf{T-Flag}. T-Flag is an attempt to improve on the previously mentioned Flag \cite{lin2010spate}. Improvements include twice the number of coloured blocks with a choice of eight colourblind-proof colours.

\textbf{Flag-Ext}. Flag Extended is the proposed scheme assessed in the paper by \textbf{H. Hsiao \textit{et al.}}\cite{hsiao2009study}. Flag extended aims again, to improve on flag by reducing the number of blocks while adding 8 possible shapes.

Figure \ref{fig:flag} contains an example of each version of Flag discussed.

\begin{figure}[h!]
    \centering
    \minipage{0.32\textwidth}
        \includegraphics[width=.9\textwidth]{flag/Flag.png}
        \caption{Flag}
    \endminipage
    \minipage{0.32\textwidth}
        \includegraphics[width=.9\textwidth]{flag/T-Flag.png}
        \caption{T-Flag}
    \endminipage
    \minipage{0.32\textwidth}
        \includegraphics[width=.9\textwidth]{flag/Flag-ext.png}
        \caption{Flag Ext}
    \endminipage
    \caption{Variations of Flag}
    \label{fig:flag}
\end{figure}

\newpage

\textbf{OpenSSH}. OpenSSH visual host key is an ASCII based graphical scheme used in MiTM detection when connecting to SSH servers. The image is generated by movement of the marker across a grid where every 2-bits of the generated MD5 has defines the diagonal direction of movement. For example, if the first 2-bits are \verb|11| the marker will move South East. The different characters are determined by the number of times the marker is present in that index, for example, if the maker is present on a square 4 times the character will be an equals sign (\verb|=|).

\textbf{Unicorns}\footnote{https://unicornify.pictures/}. Unicorns can be considered an example of an ``Avatar'' graphical scheme. The provided hash changes aspects of the image such as background colour, horn length and colour of the Unicorn's mane.

\textbf{Vash}\footnote{https://github.com/thevash/vash}. Vash is similar to that of unicorns where the hash determines certain characteristics of the image. Vash claims to have $\approx 5,438$ bits of entropy.

\begin{figure}[h!]
    \centering
    \minipage{0.32\textwidth}
        \includegraphics[width=.9\textwidth]{graphical/openssh.png}
        \caption{OpenSSH Visual Host Key}
    \endminipage
    \minipage{0.32\textwidth}
        \includegraphics[width=.9\textwidth]{graphical/unicorn.png}
        \caption{Unicorn}
    \endminipage
    \minipage{0.32\textwidth}
        \includegraphics[width=.9\textwidth]{graphical/vash.png}
        \caption{Vash}
    \endminipage
    \caption{Examples of alternative graphical encoding schemes}
\end{figure}

Research in this area has taken these various encoding schemes and compared their fallibility to impersonated attacks. The main elements of comparison have been the ``accuracy of attack detection" and ``time to compare." In this paper, these are considered the metrics of ``security'' and ``usability,'' respectively.

\section{Definition of terms}
Throughout the paper, various terms are used to define elements of the authentication ceremony relevant to fingerprints verification. Explicit descriptions are, therefore, stated in this section. These are used to remove ambiguity and underlying associations readers may have for the words chosen.
\\\\
\textbf{\textit{Method of comparison}} - is the way a user assesses the encoding scheme. The most common example and the one used as default is ``Compare-and-Confirm'' (CaC). CaC, as the name suggests, is the \textit{comparison} of two fingerprints on different devices or mediums that are then \textit{confirmed}. The following sections explain alternative methods of comparisons.

\todo{Phase this section out - make sure to remove from the introduction layout of the section}

@@

\subsubsection*{Encoding scheme performance}
Results 
from the literature consistently show the effectiveness of language-based encodings such as Words or Sentences with accuracies ranging up from 94\% \cite{dechand2016empirical}\cite{tan2017can}\cite{kainda2009usability}. In all cases, these were the best schemes from the sets assessed. The exception to this is the work performed by \textbf{H. Hsiao \textit{et al.}}\cite{hsiao2009study} in 2009 with Words achieving an abnormal accuracy of 63\%.
\\\\
Aside from textual representations were graphical schemes. As explained in Section \ref{sec:encodingSchemes} the main graphical schemes assessed by the literature were: Random Art, Flag, T-Flag, Vash, OpenSSH Visual Host Key and Unicorns. These schemes had mixed accuracy with ranges as large as 50\% - 94\% in work by \textbf{Hsiao \textit{et al.}}\cite{hsiao2009study}. The only other paper assessing graphical representations was the work of \textbf{Tan \textit{et al.}}\cite{tan2017can} where they also achieved mixed results with accuracies ranging from 46\% to 90\%.
\\
\begin{table}[h!]
    \makebox[\textwidth][c]{
        \input{tables/background/fingerprint_times.tex}
    }%
    \caption{Average comparison time (seconds) for the encoding schemes}
    \label{tab:times}
\end{table}
\\
In terms of usability of graphical schemes, the literature concurred on their high usability. The comparison speed of these schemes were all among the quickest (See Table \ref{tab:times} for an overview of timings). Other work also indirectly agreed with graphical encodings having significantly quicker comparison times compared to non-graphical schemes \cite{dechand2016empirical, kainda2009usability}.
In terms of research into the performance of graphical schemes, the literature does not contain an extensive review with literature only containing two papers. There is also no overlap in the schemes assessed with each paper reviewing a unique set. This area is, therefore, a promising candidate for further research.

One unique paper in this research area was the work by \textbf{M. Shirvanian, N. Saxena and J. J. George}\cite{shirvanian2017pitfalls} produced in the context of secure messaging pairing. This paper was unique for several reasons. First was to consideration for ``remote-vs-proximity'' pairing where this is the first consideration of this aspect found in the literature. There is room for further research to compare encoding schemes in the context of ``remote-vs-proximity.'' Another unique aspect was the end-to-end encryption context of the study.

@@

Alongside this, it was shown in work by \textbf{H. Hsiao \textit{et al.}}\cite{hsiao2009study} that age and gender do not affect the accuracy of the scheme. However, younger participants were considerably faster. Furthermore, findings also showed that language comprehension helped in discerning small differences between schemes encoded in Chinese, Japanese, Korean, or English. Subsequently, knowledge of the language did not assist in differentiating more significant changes in the schemes as these had high accuracy regardless. These were unique considerations. Further work could aim to corroborate these conclusions.

\begin{table}[h!]
    \makebox[\textwidth][c]{
        \input{tables/background/fingerprint_comparison.tex}
    }%
    \caption{Overall accuracy of correct comparison for the encoding schemes assessed}
    \label{tab:results}
\end{table}

Tables \ref{tab:times} \& \ref{tab:results} contain the accuracy results of all papers assessed; this is to aid in visual comparison. Each paper used a different metric for measuring accuracy. Therefore, all results have been translated into ``overall accuracy.''. For example, if a paper presented "security failures" or "attack success rate" of 5\% the overall accuracy of the scheme is 95\%.

@@

The starkest limitations of the literature are the range of participants and encoding entropy. The worst studies tested only 22-bits of entropy. This lack of entropy makes it challenging to compare results directly. One of the papers this most effects is the early work by \textbf{H. Hsiao \textit{et al.}}\cite{hsiao2009study}  where their highest entropy is 28-bits. This level of entropy was inadequate even at the time of publication. There is an attempt by the authors' to address this issue in the later stages of the paper where they write: \textit{``[...] increasing entropy is not a solution because it sacrifices usability and accuracy. With more entropy, representations will contain longer sequences of characters or more minute details which will lead to increased time and errors during comparisons.''}. This statement is backed up with no empirical evidence, and the authors fail to consider how the low entropy would affect the overall security of the schemes.

@@

One consideration of note is that made by \textbf{S. Dechand \textit{et al.}}\cite{dechand2016empirical} where they briefly consider medical conditions such as ADHD and reading or visual disorders and the way they affect the comparison's effectiveness. The results highlight a slight reduction in overall accuracy, however, due to their small sample size, the results cannot be considered conclusive. This aspect is unique to all literature assessed. Further work would be required to produce conclusive results. Therefore, highlighting a gap in the literature.

One glaring issue with the demographical health of the study by \textbf{E. Uzun, K. Karvonen} and \textbf{N. Asokan}\cite{uzun2007usability} is the use of two entirely different groups of participants. Not only were their demographics different, but they were from different countries (America and Finland). Different cultures contain inherent biases and assumptions. Changes were made pragmatically regarding the results from the first round of 40 participants. These were then re-assessed, alongside the direct comparison of results. This aspect casts vast doubts on the validity of the results with no consideration made by the authors to control external factors that may affect the performance of the method of comparison.

Hexadecimal and numerical schemes were not included as encoding schemes in work by \textbf{H. Hsiao \textit{et al.}}\cite{hsiao2009study} (some of the most widespread encoding schemes). The justification was the schemes similarities to Base32  alongside "well-known deficiencies". This point was provided with no further justification or quantified. Moreover, it is also inconsistent with available research, for example, in \textit{``Empirical Study of Textual Key-Fingerprint Representations''}\cite{dechand2016empirical} it was shown numerical representations performed significantly better than that of Base32.

\begin{table}[h!]
    \makebox[\textwidth][c]{
        \input{tables/background/fingerprint_attacker_stren.tex}
    }%
    \caption{Paper attribute comparison}
    \label{tab:attacker}
\end{table}

@@

\subsubsection*{Conclusion}
Overall, this review has identified several key areas suitable for further work.
The first is the performance assessment of graphical encoding schemes. Recreation of pre-existing results from \cite{hsiao2009study}\cite{tan2017can} is required to corroborate current conclusions and validate results.
Another gap in the research is the consideration into the utilisation of encoding schemes in realistic conditions, i.e. ``remove vs proximity." This topic was initially covered by \cite{shirvanian2017pitfalls}, but their scope was limited. Further work, therefore, could increase the scope and touch upon a large number of schemes in these settings. The final aspect for further work is the limited consideration into the feasibility of attacks on encoding schemes. All of the papers assessed simulated attacks and had minimal consideration for the execution of these attacks. Therefore, further work could delve
into the implementation and feasibility of such attacks.

@@

Some of the oldest preliminary work into visual encoding schemes is by \textbf{A. Perrig} and \textbf{D. Song}\cite{perrig1999hash} in the creation of their scheme ``Random Art" in 1999. The motivation for creating such a scheme was the perceived flaws in the ways humans verify and compare written information. As mentioned in previous sections, visual encoding schemes have been shown to have mixed success, with low security being one of their most alarming flaws. This research laid the foundation for further work in analysing the security of visual encoding schemes.

Further research into the creation of unique visual hash schemes has been performed by \textbf{C. Ellison} and \textbf{S. Dohrmann}\cite{ellison2003public} (Flag), \textbf{Yue-Hsun Lin \textit{et al.}}\cite{lin2010spate} (T-Flag) and work by \textbf{M.  Olembo \textit{et al.}}\cite{olembo2013developing}. Each publication has provided a new way  to represent a key fingerprint visually. Alongside the academic literature, there are informally presented methods of visual fingerprints such as Unicorns\footnote{https://unicornify.pictures/} and Robots\footnote{https://github.com/e1ven/Robohash}. This list is by no means exhaustive but is used to depict the amount of research and work invested into graphical hash representations.

@@

Overall the paper provides a foundation for formalising the creation of effective wordlists. A limitation is the lack of empirical data gathered on the performance. However, this was later evaluated in work by \textbf{S. Dechand \textit{et al.}}\cite{dechand2016empirical} and shown to be an effective encoding scheme.

Other research of note is work by \textbf{M. Goodrich \textit{et al.}}\cite{goodrich2006loud} called \textit{Loud and Clear: Human-Verifiable Authentication Based on Audio}. As the name suggests, the authors were researching ways to improve current methods of secure device pairing. The unique aspect of this work is the use of a Text-to-Speech system reading out syntactically correct English sentences. MadLibs\footnote{https://en.wikipedia.org/wiki/Mad\_Libs} was the base for these sentences. MadLibs is where static placeholders are replaced with a list potential words.\\
This work is an extension to work performed by \textbf{Juola} and \textbf{Zimmermann}\cite{juola1996whole} as they aimed to emulate the techniques used in PGPfone. The paper's findings are limited by the lack of empirical data backing up claims made by the author the performance of the system, and security is only theoretically assessed.

Tanet alAside from this research, there have been further informal implementations of fingerprint encodings. The first being by \textbf{Michael Rogers}\footnote{https://github.com/akwizgran/basic-english}. Rogers' implementation is a program designed to map fingerprints to pseudo-random poems. This implementation was again, empirically evaluated by \textbf{S. Dechand \textit{et al.}}\cite{dechand2016empirical}. Ealier work by \textbf{N. Haller} with the S/KEY\cite{haller1995s} shows the implementation of a system designed to represent a hash as a series of six short words. However, this system is designed for a one-time-password purpose and only provides word mappings for basic human usability of the password and not within a fingerprint verification context. Therefore, the wordlist has not been designed with pronounceability in mind.

@@

\subsubsection*{Conclusion}

In conclusion to this topic, the current research has primarily focused on the research and creation of visual representations. Research for textual fingerprints is fragmented and incomplete with work \textbf{Juola} and \textbf{Zimmermann} 
\cite{juola1996whole} and \textbf{M. Goodrich \textit{et al.}}\cite{goodrich2006loud} providing meaningful research to build upon in terms of word and sentence based encodings. The fragmentation of this research leaves room for further work into this topic area. Alongside this, findings from the previous section's research show that human language based encodings provided the best usability and, therefore, should be a target for further research looking to improve upon their security and usability.

@@

The only other relevant research on this topic is the work by \textbf{M Shirvanian and N. Saxena}\cite{shirvanian2014wiretapping} 
and their paper: \textit{``Wiretapping via Mimicry: Short 
Voice Imitation Man-in-the-Middle Attacks on Crypto 
Phones''}. Further research in the area of ``human voice impersonation'' has received lots of attention \cite{mukhopadhyay2015all}\cite{chen2017you}\cite{wu2015spoofing}. This paper was chosen over other alternatives due to its specific use of encoding schemes in its evaluation.

In this paper, the authors develop a way to 
impersonate users when authenticating 
Short-authentication-Strings (SAS) in the pairing of 
Crypto-phones. To achieve this impersonating they propose 
two methods: ``Short voice re-ordering attack'' where an 
arbitrary SAS string is recreated by re-ordering snipped 
obtained from eavesdropping a previous connection
and ``Short voice morphing attacks'' whereby the use of 
previously eavesdropped audio snippets the attacker can
morph their voice to match that of the victim. With 
these methods, they aimed to attack encodings of Numbers, 
PGP word list (previously discussed work by \textbf{Juola} and 
\textbf{Zimmermann} \cite{juola1996whole}) and MadLib (\textbf{M. Goodrich 
\textit{et al.}}\cite{goodrich2006loud} work also 
previously discussed). The effectiveness of these attacks 
was evaluated with a study involving 30 participants.

@@

\section{Overall Summary}
In summary of the literature gaps areas discovered; encoding scheme performance requires further work in the performance of all graphical schemes to back up the results made by other work. Furthermore, there is a gap with the assessment of encoding schemes in the context of "remote-vs-proximity" first proposed by \textbf{M. Shirvanian, N. Saxena} and \textbf{J. J. George}\cite{shirvanian2017pitfalls} as it would arguably provide a better simulation of real-world scenarios.

In the context of human demographics and the way they alter the performance of the schemes; further research is required into how the fluency of languages affects authentication ceremony. This further work would be a continuation of the initial work by \textbf{H. Hsiao \textit{et al.}}\cite{hsiao2009study} and could allow the creation of schemes that are better suited to specific languages.
Alongside this, there is a research gap in the investigation of how mental impairments affect the performance. This aspect is useful due to the number of potential users being affected due to systems not being designed with them in mind. For example, 7\% of the population have been identified as having dyslexic tendencies\cite{peterson2012developmental}. This fact means that with a UK population of around 63,000,000\footnote{From the 2011 Census collated by the Office for National Statistics}  4,410,000 people could benefit from encoding schemes designed to work well with dyslexia. Dyslexia, however, is only one of many impairments that could affect a scheme's performance, thus further contributing to the substantial effect further work could have on this area.
